---
title: "gun_deaths"
format: html
---

```{r warning=FALSE, message=FALSE}
library(tidyverse)
gun_data <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/guns-data/master/full_data.csv")

## glimpse(gun_data)

## head(gun_data)
```

```{r}
gun_data %>%
count(intent) %>%
ggplot(aes(x = reorder(intent, -n), y = n, fill = intent)) +
geom_col(show.legend = FALSE) +
labs(title = "Gun Deaths by Intent (2018)",
x = "Intent",
y = "Number of Deaths") +
theme_minimal() +
theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

This plot shows that suicides account for the vast majority of gun deaths in 2018, followed by homicides. This highlights how prevention efforts could focus more on mental health resources and suicide awareness campaigns, especially targeting high-risk groups.

To address the client’s request for identifying seasonal emphasis areas, this chart examines monthly variations.

```{r}

gun_data %>%
count(month) %>%
ggplot(aes(x = month, y = n, group = 1)) +
geom_line(color = "steelblue", size = 1) +
geom_point(color = "steelblue", size = 2) +
labs(title = "Gun Deaths by Month (2018)",
x = "Month",
y = "Number of Deaths") +
theme_minimal()
```

There appears to be a slight increase in gun deaths during the summer months (June–August), suggesting that awareness campaigns focusing on safe conflict resolution and stress management might be most effective during this time. Winter months generally show fewer deaths, possibly due to seasonal behaviors or reduced outdoor activity.

Next, we group individuals by age to understand which life stages are most affected.

```{r}

gun_data %>%
mutate(age_group = cut(age, breaks = c(0, 20, 40, 60, 80, 100),
labels = c("0–19", "20–39", "40–59", "60–79", "80+"))) %>%
count(age_group) %>%
ggplot(aes(x = age_group, y = n, fill = age_group)) +
geom_col(show.legend = FALSE) +
labs(title = "Gun Deaths by Age Group (2018)",
x = "Age Group",
y = "Number of Deaths") +
theme_minimal()
```

Gun deaths are most common among individuals aged 20–39, followed by those aged 40–59. Younger adults appear to be at the highest risk, indicating that outreach efforts and educational campaigns targeting this demographic could have the greatest impact.

Finally, this chart explores gender differences in types of gun deaths.

```{r}

gun_data %>%
count(sex, intent) %>%
ggplot(aes(x = sex, y = n, fill = intent)) +
geom_col(position = "dodge") +
labs(title = "Gun Deaths by Gender and Intent (2018)",
x = "Gender",
y = "Number of Deaths",
fill = "Intent") +
theme_minimal()
```

Males represent the overwhelming majority of gun deaths, particularly in suicides. Female deaths occur far less frequently but are more likely to be homicides. This suggests that messaging for male audiences could focus on mental health and suicide prevention, while female-focused campaigns could emphasize domestic violence prevention and safety resources.
