---
title: "string_notes"
format: html
---

```{r}
library(pacman)
pacman::p_load(tidyverse, stringr, stringi, rio)

tibble(sentence = sentences[1:4]) %>% 
  mutate(
    n_the = str_count(sentence, pattern = "the |The"),
    n_theB = str_count(tolower(sentence), pattern = "the"),
    n_lowercase = str_count(sentence, pattern = "[:lower:]")
  )
```

```{r}
chars <- readr::read_lines('https://byuistats.github.io/M335/data/randomletters.txt')

chars_num <- readr::read_lines("https://byuistats.github.io/M335/data/randomletters_wnumbers.txt")
```

Task 1:

1.  What is the length of the randomletters string? ﻿﻿
2.  Remove all the 'e' and 'a' letters and then tell me how long the string is. and 'a' letters ﻿﻿
3.  How many times is the name 'jim' in the string?
4.  ﻿﻿Tell me which character locations have three "a"'s in a row. ﻿﻿
5.  Show all the sequences with 3 of the same letter in a row. (hardest one?) ﻿﻿
6.  Split the string so that instead of a stzing of length 70000, it is a vector of character strings that eacl have length 1.{r} str_length(chars)

```{r}
str_extract_all(chars, "(.)\\1{2}")[[1]]

```
