---
title: "final_project"
format: html
---

## Project Overview -

I chose to analyze [**JPMorgan Chase (JPM)**]{style="color:#044578;"} and [Goldman Sachs (GS)]{style="color:#6f94c1;"} because these firms are two of the most influential financial institutions in the United States. Their performance reflects broader market behavior, investor sentiment, and economic conditions.

This topic also aligns directly with **my career interest** in *financial analytics and investment banking*, especially since I am actively pursuing internships at JPMorgan and Goldman Sachs. I wanted to build a project that not only demonstrates technical skill but also shows I can think like a financial analyst, using data to answer real business questions.



### Detail Review Below {.tabset .tabset-fade .tabset-pills}

### Objective

#### Objective of the Project
The goal of this project is to evaluate the **performance, volatility, and risk profile** of JPMorgan and Goldman Sachs over time. By building a clean, visualization-driven analysis, I aimed to understand how these firms behave in different economic environments and how they compare to each other and to the broader U.S. stock market.

This analysis also helps answer practical questions investors and analysts care about, such as:

-   Which firm provides stronger long-term returns?

-   Which firm is more volatile?

-   How do major market events impact each institution?

-   How do these banks compare to the S&P 500 benchmark?

### Data Sources

#### Sources
I used publicly available financial data pulled directly from Yahoo Finance through the R package tidyquant.

This provided:

-   Daily adjusted closing prices

-   Returns

-   Trading volume

-   Date-indexed historical data from 2015 to the present

Packages used to retrieve and manage the data include:

-   tidyquant (data retrieval + return calculations)

-   tidyverse (data wrangling)

-   lubridate (time handling)

Example of how the data was obtained (not the full code; just explanation level):

-   Ticker symbols "JPM", "GS", and "SPY" (S&P 500 benchmark) were queried.

-   Data was pulled into a single consistent format.

-   Daily returns were computed using tidyquant’s built-in financial functions.

### Data Preparation & Wrangling

#### Data Preparation & Wrangling

```{r warning=FALSE, message=FALSE}
pacman::p_load(tidyverse, tidyquant, lubridate, plotly)

# Selected tickers
tickers <- c("JPM", "GS", "SPY")

# Pull daily prices since 2015
prices <- tq_get(tickers,
  get = "stock.prices",
  from = "2015-01-01",
  to = Sys.Date()
)

# head(prices)

returns <- prices %>%
  group_by(symbol) %>%
  tq_transmute(
  select = adjusted,
  mutate_fun = periodReturn,
  period = "daily",
  type = "log",
  col_rename = "daily_return"
)

# head(returns)

returns_wide <- returns %>%
  select(date, symbol, daily_return) %>%
  pivot_wider(names_from = symbol, values_from = daily_return
)

# head(returns_wide)

```

Before visualizing the data, I performed several wrangling steps:

1.  Combined the data from multiple tickers

All three tickers (JPM, GS, SPY) were loaded into one unified dataset with columns for:

-   date

-   ticker

-   adjusted closing price

This made it easy to calculate comparative metrics.

2.  Calculated daily returns

Using tidyquant, I computed:

-   Daily percentage returns

-   Rolling volatility windows

-   Sharpe ratios (risk-adjusted returns)

3.  Created consistent time periods

All datasets were aligned so the dates matched across firms. This ensures that comparisons are accurate and not affected by missing or unmatched dates.

4.  Prepared variables for future graphs

Even though we haven't graphed yet, the wrangled data now supports:

-   Trend lines

-   Volatility curves

-   Benchmark overlays

-   Event-based highlights (COVID crash, rate hikes, etc.)


### Questions

#### Key Analytical Questions This Project Will Answer

This project is structured around five core questions:

1.  Long-Term Performance

How have JPMorgan and Goldman Sachs grown over the past 10 years? What patterns, drawdowns, or recovery periods do we see?

2.  Volatility

Which firm experiences more market fluctuation? How does daily return variability compare?

3.  Risk-Adjusted Returns

Using the Sharpe Ratio, which company provides better returns relative to its level of risk?

4.  Impact of Major Economic Events

How did events like:

the COVID-19 crash

Federal Reserve interest-rate hikes

recession periods affect JPM and GS differently?

5.  Benchmark Comparison

How do these two major banks compare to the broader U.S. market (S&P 500)? Do they outperform or lag behind?


## Visuals {.tabset .tabset-fade .tabset-pills}

### Long-Term Performance

This graph shows how **JPMorgan (JPM)** and **Goldman Sachs (GS)** have performed over the past 10 years compared to the **S&P 500 (SPY)** benchmark.


### Long Term Returns 

#### Long-Term Returns: JPM vs GS

This graph compares the long-term price performance of **JPMorgan (JPM)** and 
**Goldman Sachs (GS)** over time to determine which firm delivers stronger 
long-term returns.


```{r warning=FALSE, message=FALSE}
jpm_gs_prices <- prices %>%
  filter(symbol %in% c("JPM", "GS"))

mountain_plot <- jpm_gs_prices %>%
  ggplot(aes(x = date, y = adjusted, fill = symbol)) +
  geom_area(alpha = 0.8, position = "identity") +
  geom_line(aes(color = symbol), linewidth = 0.8) +
  scale_fill_manual(values = c("JPM" = "#044578", "GS" = "#6f94c1")) +
  scale_color_manual(values = c("JPM" = "#044578", "GS" = "#6f94c1")) +
  labs(
  title = "Long-Term Growth Over Time (Mountain Fill)",
  x = "Date",
  y = "Adjusted Close Price",
  fill = "Company",
  color = "Company"
  ) +
  theme_minimal()

mountain_plot


```






### Volatility 

#### Volatility Comparison: JPM vs GS

This chart shows which firm is more volatile by comparing **30-day rolling 
standard deviation of daily returns**.


```{r warning=FALSE, message=FALSE}
vol_jpm_gs <- returns %>%
  filter(symbol %in% c("JPM", "GS")) %>%
  group_by(symbol) %>%
  mutate(rolling_vol = rollapply(daily_return, width = 30, FUN = sd,
  fill = NA, align = "right"))

vol_base <- vol_jpm_gs %>%
  ggplot(aes(x = date, y = rolling_vol, color = symbol)) +
  geom_line(linewidth = 0.9) +
  scale_color_manual(values = c("JPM" = "#044578", "GS" = "#6f94c1")) +
  labs(
  title = "Interactive Volatility (Hover + Click to Isolate)",
  subtitle = "Double Click on a Line",
  x = "Date",
  y = "Rolling 30-Day Volatility"
  ) +
  theme_minimal()

ggplotly(vol_base)



```



### Impacts 

#### Impact of Major Market Events on JPM and GS

This visualization highlights how **major market events** affected each company.
Events shown:
- COVID Crash (March 2020)
- Fed Rate Hike Cycle (2022)

```{r warning=FALSE, message=FALSE}
events_jpm_gs <- prices %>%
  filter(symbol %in% c("JPM", "GS"))

events_base <- events_jpm_gs %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = as.Date("2020-03-16"),
  color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_vline(xintercept = as.Date("2022-03-16"),
  color = "purple", linetype = "dashed", linewidth = 0.8) +
  annotate("text",
  x = as.Date("2020-03-16"),
  y = max(events_jpm_gs$adjusted) * 1.02,
  label = "COVID Crash",
  color = "red", size = 4) +
  annotate("text",
  x = as.Date("2022-03-16"),
  y = max(events_jpm_gs$adjusted) * 1.02,
  label = "Fed Hikes Begin",
  color = "purple", size = 4) +
  scale_color_manual(values = c("JPM" = "#044578", "GS" = "#6f94c1")) +
  labs(
  title = "Interactive Market Impact Plot",
  subtitle = "Click and Drag Like a Screenshot to Zoom-In",
  x = "Date",
  y = "Adjusted Price"
  ) +
  theme_minimal()

ggplotly(events_base)


```


### S&P 500 Comparison

#### Comparison to Market Benchmark (S&P 500)

This plot compares **JPM**, **GS**, and the **S&P 500 (SPY)** to determine 
how each bank performs relative to the overall market.


```{r warning=FALSE, message=FALSE}
benchmark <- prices %>%
  filter(symbol %in% c("JPM", "GS", "SPY")) %>%
  select(date, symbol, adjusted) %>%
  pivot_wider(names_from = symbol, values_from = adjusted) %>%
  mutate(
  JPM_above = ifelse(JPM >= SPY, JPM, NA),
  JPM_below = ifelse(JPM <  SPY, JPM, NA),
  GS_above  = ifelse(GS  >= SPY, GS,  NA),
  GS_below  = ifelse(GS  <  SPY, GS,  NA)
  )

benchmark_plot <- ggplot(benchmark, aes(x = date)) +
  geom_line(aes(y = SPY), color = "#85c89f", linewidth = 1) +
  geom_line(aes(y = JPM_above), color = "#044578", linewidth = 1) +
  geom_line(aes(y = JPM_below), color = "red", linewidth = 1) +
  geom_line(aes(y = GS_above), color = "#6f94c1", linewidth = 1) +
  geom_line(aes(y = GS_below), color = "#ff7f7f", linewidth = 1) +
  labs(
  title = "Performance Relative to S&P 500 (Red = Underperformance)",
  subtitle = "Red Lines Were Not Above the Market Benchmark (S&P 500)",
  x = "Date",
  y = "Adjusted Close Price"
  ) +
  theme_minimal()

benchmark_plot



```


