---
title: "string_task"
format: html
---


Question 1 (Use the `readr::read_lines()` function to read in each: [string—randomletters.txt](https://byuistats.github.io/M335/data/randomletters.txt) and [randomletters_wnumbers.txt.](https://byuistats.github.io/M335/data/randomletters_wnumbers.txt))

```{r}
pacman::p_load(tidyverse, stringr, stringi, rio)




```

Question 2 (With the `randomletters.txt` file, pull out every 1700 letter (for example, 1, 1700, 3400, 5100, …) and find the quote that is hidden—the quote ends with a period.)

```{r}
chars <- readr::read_file('https://byuistats.github.io/M335/data/randomletters.txt')
inculde_all <- paste(chars, collapse = "")
break_string <- str_split(inculde_all, "")[[1]]

position <- c(1,seq(0, length(break_string), by = 1700))

break_string[position]


```

Question 3 (With the `randomletters_wnumbers.txt` file, find all the numbers hidden, and convert those numbers to letters using the letters order in the alphabet to decipher the message. For example, a 1=a, 2=b,…, 26=z (Hint: the message starts with “experts”).)

```{r}
chars_num <- readr::read_lines("https://byuistats.github.io/M335/data/randomletters_wnumbers.txt")

num_words <- str_extract_all(chars_num, "[:digit:]+")[[1]]
num_words <- as.integer(num_words)
nums <- letters[num_words]
paste(nums, collapse = "")
```

Question 4\
With the `randomletters.txt` file, remove all the spaces and periods from the string then find the longest sequence of vowels.

```{r}
cleaned <- str_remove_all(chars, "[ \\.]")
vowel_sequences <- str_extract_all(cleaned, "[aeiouy]+")[[1]]
longest_vowel_sequence <- vowel_sequences[which.max(nchar(vowel_sequences))]
longest_vowel_sequence
```

